---
title: "7 - intercations"
author: "Aske Qvist"
date: "26 March 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 7 - interactions

Loading data
```{r}
# Packages
library(pacman)
p_load(rethinking, brms, tidyverse, patchwork)

# Data
data(rugged)
d <- rugged

detach(package:rethinking, unload = T)

```

Make data subsets
```{r}

# make log version of outcome
d <- d %>% mutate(
  log_gdp = log(rgdppc_2000)
)

# extract countries with GDP data
dd <- d %>% filter(complete.cases(rgdppc_2000))

# split countries into Africa and not-Africa
d.A1 <- dd %>% filter(cont_africa == 1) # Afrika
d.A0 <- dd %>% filter(cont_africa == 0) # not Afrika

```

Fit models to the two data subsets
```{r}

b7.1 <-
  brm(data = d.A1, family = gaussian, # Afrikan data
      log_gdp ~ 1 + rugged,
      prior = c(prior(normal(8, 100), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(uniform(0, 10), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 7)

b7.2 <-
  update(b7.1, # Not Afrikan data
         newdata = d.A0)

```

Plot posterior predictions
```{r}
africa <- d.A1 %>%
  ggplot(aes(x = rugged, y = log_gdp)) +
  stat_smooth(method = "lm", fullrange = T, size = 1/2,
              color = "firebrick4", fill = "firebrick", alpha = 1/5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 1/2) +
  scale_x_continuous("Rugged", limits = c(0, 55)) +
  coord_cartesian(xlim = 0:7, ylim = 6:10) +
  ylab("Log of GDP") +
  theme_bw() + ggtitle("Africa") 
  
not_africa <- d.A0 %>%
  ggplot(aes(x = rugged, y = log_gdp)) +
  stat_smooth(method = "lm", fullrange = T, size = 1/2,
              color = "firebrick4", fill = "firebrick", alpha = 1/5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 1/2) +
  scale_x_continuous("Rugged", limits = c(0, 55)) +
  coord_cartesian(xlim = 0:7, ylim = 6:10) +
  ylab("Log of GDP") +
  theme_bw() + ggtitle("Not Africa")
  

not_africa + africa
# slope is positive within africa but negative outside africa


```

It is better if we keep the data in the same data set. Multilevel models borrow data from the different categories in order to improve estimates in all categories - and many other reasons. But even if we added the dummy variable cont_africa, this will not solve the problem - as a normal linear regression does not account for different slopes for the categories of the dummy.

# Comapre the models with/without the dummy cont_africa as predictor ( but no interaction effect)
```{r}
# Start with the model I made before (without the dummy cont_africa) - just with all the data
b7.3 <-
  update(b7.1, newdata = dd) # all data

# Now the same but added is the dummy as a predictor (in the formula)
b7.4 <-
  update(b7.3,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged + cont_africa) 

# compute information criteria for both
b7.3 <- add_criterion(b7.3, c("loo", "waic"))
b7.4 <- add_criterion(b7.4, c("loo", "waic"))


# compare IC
loo_compare(b7.3, b7.4,
            criterion = "waic")

loo_compare(b7.3, b7.4,
            criterion = "loo")
# loo and waic agree - the model with the dummy fits teh data way better

# compute weight
model_weights(b7.3, b7.4,
              weights = "waic") %>% 
  round(digits = 3)
# The model with the dummy has all the weight :)


```


# adding a linear interaction
We want the realtionship between Y (GDP) and R (rugged) to vary as a function of A (dummy varible)

(the model comparisons do not work) add_criterion() and model_weigths() when using brms.
```{r}

# Define new model with interaction effect 
b7.5 <-
  update(b7.4,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged*cont_africa) # syntax as lmer()

b7.5b <-
  update(b7.4,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged + cont_africa + rugged:cont_africa) 

# add IC 
b7.5 <- add_criterion(b7.5,c("loo","waic"))

# compare all 3 models
l <- loo_compare(b7.3, b7.4, b7.5, criterion = "waic")

print(l, simplify = F)

# weight
model_weights(b7.3, b7.4, b7.5,
              weights = "loo") %>% 
  round(digits = 3)

b7.5$loo # This is not working!

```








